---
title: "HW4 - Math 181B"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Question 1.
  $$
  Brown: n_1=45, p_1=0.3 \\
  Yellow: n_2 = 34, p_2=0.2\\
  Red: n_3=32, p_3=0.2\\
  Orange: n_4=15, p_4=0.1\\
  Blue: n_5=13, p_5=0.1\\
  Green: n_6=14, p_6=0.1\\
  $$
  
##### a) What is your null hypothesis and alternative hypothesis?
  $$
  H_0: p_1 = p_{10}, p_2 = p_{20}, p_3 = p_{30}, p_4 = p_{40}, p_5 = p_{50}, p_6 = p_{60} \\
  H_1: p_j != p_{j0}, j \in [1,2,3,4,5,6]
  $$
  
##### b) What is the total number of the observations n?
```{r}
X <- c(45, 34, 32, 15, 13, 14)
n <- sum(X)
print(n)
```

##### c) What is the value of the χ2 test statistic?
  _x^2 test statistics is:_
  $$
  m = 6 \\
  D = \sum_{j=1}^m \frac{(X_j - np_j)^2}{np_j} \\
  df = m - 1
  $$
```{r}
m <- 6
p0 <- c(0.3, 0.2, 0.2, 0.1, 0.1, 0.1)
d0 <- rep(0, m)
for (j in 1:m) {
  d0[j] <- (X[j] - n*p0[j])^2 / (n*p0[j])
}
d <- sum(d0)
print(d)
```

##### d) What is the p-value?
```{r}
p.value <- pchisq(d, df = m-1, lower.tail = FALSE)
print(p.value)
alpha <- 0.05
p.value <= alpha
```

##### e) What is your conclusion?
_since the p-value is greater than the level of significant 0.05, we fail to reject H0._


## Question 2.
```{r}
X = c(251, 264, 234, 283, 226, 244, 269, 241, 276, 274, 263, 243, 254, 276, 241, 232, 260, 248, 284, 253, 265, 235, 259, 279, 256, 256, 254, 256, 250, 269, 240, 261, 263, 262, 259, 230, 268, 284, 259, 261, 268, 268, 264, 271, 263, 259, 294, 259, 263, 278, 267, 293, 247, 244, 250, 266, 286, 263, 274, 253, 281, 286, 266, 249, 255, 233, 245, 266, 265, 264, 257, 261, 252, 254, 266)
n0 = rep(0, 8)
for (j in 1:length(X)) {
  if ((220 <= X[j]) && (X[j] < 230)) {
    n0[1] = n0[1] + 1;
  }
  else if ((X[j] >= 230) && (X[j] < 240)) {
    n0[2] = n0[2] + 1;
  }
  else if ((X[j] >= 240) && (X[j] < 250)) {
    n0[3] = n0[3] + 1;
  }
  else if ((X[j] >= 250) && (X[j] < 260)) {
    n0[4] = n0[4] + 1;
  }
  else if ((X[j] >= 260) && (X[j] < 270)) {
    n0[5] = n0[5] + 1;
  }
  else if ((X[j] >= 270) && (X[j] < 280)) {
    n0[6] = n0[6] + 1;
  }
  else if ((X[j] >= 280) && (X[j] < 290)) {
    n0[7] = n0[7] + 1;
  }
  else if ((X[j] >= 290) && (X[j] < 300)) {
    n0[8] = n0[8] + 1;
  }
}
print(n0)

```
  $$
  Y_1: 220<=y<230; Observed \#: 1 \\
  Y_2: 230<=y<240; Observed \#: 5 \\
  Y_3: 240<=y<250; Observed \#: 10 \\
  Y_4: 250<=y<260; Observed \#: 19 \\
  Y_5: 260<=y<270; Observed \#: 25 \\
  Y_6: 270<=y<280; Observed \#: 7 \\
  Y_7: 280<=y<290; Observed \#: 6 \\
  Y_8: 290<=y<300; Observed \#: 2 \\
  $$


##### a) What is your estimated μ and σ?
```{r}
Obs <- n0
total.dur <- sum(X)
n <- 75
mean <- total.dur/n
s <- 2
print(mean)
```

  $$
  \hat{\mu} = \bar{x} = 260.16 \\
  \hat{\sigma} = s = 2 \\
  $$


##### b) What is the estimated probabilities of these classes? 

```{r}
mu.hat = mean
sd.hat = s
m <- length(Obs)
l = seq(220,300, by=10)
p0 = rep(0, m)
for (j in 2:length(l)) {
  p1 = pnorm(l[j], mean = mu.hat, sd = sd.hat, lower.tail = TRUE)
  p2 = pnorm(l[j-1], mean = mu.hat, sd = sd.hat, lower.tail = TRUE)
  p0[j-1] = p1 - p2
}
print(p0)
```



##### c) What is the value of the χ2 test statistic?
```{r}
Est <- n*p0
d0 <- (Obs - Est)^2 / Est
dhat <- sum(d0)
print(dhat)

```

##### d) What is your p-value?
```{r}
p.value = 1 - pchisq(dhat, df = m - 1 - s)
print(p.value)
alpha <- 0.1
p.value < alpha
```

##### e) What is your conclusion? 
_since the p-value is less than the level of significant, we reject the H0._ 


## Question 3.
##### a) At the α = 0.01 level of significance, is there evidence here to support the contention that birth order and delinquency are related?
**H0: birth order and delinquency are independent.** 
**H1: birth order and delinquency are not independent.**
  $$
  r = 4 \\
  c = 2 \\
  total.Delinquent = 111 \\
  total.NOT-Delinquent = 1154 \\
  alpha = 0.01 \\
  H_0: P(A and B) = independent \\
  H_1: P(A and B) = NOT-independent \\
  $$
```{r}
r = 4
c = 2
x <- matrix(c(24, 29, 35, 23, 450, 312, 211, 70), ncol = c, nrow = r)
Csum <- colSums(x)
Rsum <- rowSums(x)
n <- sum(x)
est.mat <- Rsum %*% t(Csum/n)
(d <- sum((x - est.mat)^2 / est.mat))
df <- (r-1)*(c-1)
(p.value <- 1 - pchisq(d, df))
alpha <- 0.01
p.value <= alpha
```
_since the p-value is less than the level of significant, we reject H0._



## Question 4.
##### a) Show that the coefficients (a, b) that minimize the loss function L(a, b) satisfy the following equation:
  $$
  \begin{aligned}
  a:\\ 
  \frac{dL}{da} &= \sum_{i=1}^n 2\bigl(y_i - (a + bx_i)\bigr)*(-1) \\
                &= \sum_{i=1}^n -2(y_i - a - bx_i) \\
                &= 0 \\
  -2(\sum_{i=1}^n y_i -na - \sum_{i=1}^n bx_i) &= 0 \\
  (\sum_{i=1}^n y_i -na - \sum_{i=1}^n bx_i) &= 0 \\
  -na &= \sum_{i=1}^n bx_i - \sum_{i=1}^n y_i \\
     a &= \frac{1}{n} (\sum_{i=1}^n y_i - b \sum_{i=1}^n x_i) \\
     an + b \sum_{i=1}^n x_i &= n * \frac{1}{n} (\sum_{i=1}^n y_i - b \sum_{i=1}^n x_i) + b \sum_{i=1}^n x_i \\
                             &= \sum_{i=1}^n y_i - b \sum_{i=1}^n x_i + b \sum_{i=1}^n x_i \\
     an + b \sum_{i=1}^n x_i &= \sum_{i=1}^n y_i \\
  b:\\
  \frac{dL}{db} &= \sum_{i=1}^n 2\bigl(y_i - (a + bx_i)\bigr)*(-x_i) \\
                &= \sum_{i=1}^n -2x_i(y_i - a - bx_i) \\
                &= 0 \\
  -2(\sum_{i=1}^n x_iy_i -a \sum_{i=1}^n x_i b \sum_{i=1}^n x_i^2) &= 0 \\
  (\sum_{i=1}^n x_iy_i -a \sum_{i=1}^n x_i - b \sum_{i=1}^n x_i^2) &= 0 \\
  \sum_{i=1}^n x_iy_i &= a \sum_{i=1}^n x_i + b \sum_{i=1}^n x_i^2 \\
 
  
  \end{aligned}
  $$
  

##### b) Find a, b.
  $$
  \begin{aligned}
  a &= \frac{1}{n} (\sum_{i=1}^n y_i - b \sum_{i=1}^n x_i) \\
    &= \frac{1}{n} (811 - 234*b) \\
  b &= \frac{1}{\sum_{i=1}^n x_i^2} (\sum_{i=1}^n x_iy_i - a \sum_{i=1}^n x_i) \\
    &= \frac{1}{10144} (24628 - 234*a) \\
    &= \frac{1}{10144} \bigl(24628 - \frac{234}{n} (811 - 234*b)\bigr)\\
  b(1 - \frac{54756}{10144n}) &= \frac{24628}{10144} - \frac{189774}{10144n} \\
  b &= (\frac{24628}{10144} - \frac{189774}{10144n}) * (\frac{10144n}{10144n - 54756}) \\
    &= \frac{24628n - 189774}{10144n - 54756} \\
  a &= \frac{1}{n} \bigl(811 - 234*(\frac{24628n - 189774}{10144n - 54756})\bigr) \\
    &= \frac{811}{n} - \frac{234}{n} *(\frac{24628n}{10144n - 54756} - 
    \frac{189774}{10144n - 54756}) \\
    &= \frac{811}{n} - \frac{5762952}{10144n-54756} + \frac{44407116}{10144n^2 -54756n} \\
  \end{aligned}
  $$
