---
title: "HW7 -Math 181B"
output:
  pdf_document: default
  html_document: default
---

## Question 1. 
  $$
  Y_{ij} - Possion \space distribution \\
  mean = \mu_j \\
  Var(Y_{ij}) = E(Y_{ij}) = \mu_j \\
  development = 2, sheets = 6 \\
  $$
##### a) Draw a boxplot for the number of visible flaws with the developers organized in groups.
```{r}
k = 2
i = 6
set.seed(181)
(origData = matrix(c(1,4,5,6,3,7,8,6,4,9,11,10), nrow = i, ncol = k))
apply(origData, 2, var)

boxplot(origData, main = "Orginal Data")
```

##### b) Assume the number of flaws on a given negative is a Poisson random variable. Make an appropriate data transformation.
  $$
  \begin{aligned} 
  recall: Var(Y_{ij}) &= g(\mu_j) \\
  A(x) &= b * \int \frac{1}{\sqrt{g(x)}} dx + c \\
  Y_{ij} &\sim Poisson(\lambda_j) \\
  \mu_j &= \lambda_j \\
  Var(Y_{ij}) &= \lambda_j \\
              &= \mu_j \\
  g(\mu_j) &= \mu_j \\
  g(x) &= x \\
  A(x) &= \int \frac{1}{\sqrt{x}} dx + 0 \\
       &= 2 \sqrt{x} \\
  \end{aligned}
  $$


##### c) Draw a new boxplot for the new data.
```{r}
(scaledData = 2*sqrt(origData))
apply(scaledData, 2, var)
boxplot(scaledData, main = "Transformed Data")
```


##### d) Do the indicated analysis of variance for the transformed data.
```{r}
data.transform = data.frame(obs = c(scaledData[,1],scaledData[,2]), class = c(rep("A", 6), rep("B", 6)))
aov.fit1= aov(obs ~ class, data = data.transform)
summary(aov.fit1)
```


## Question 2.
##### a) Do the analysis of variance for the data in the table.
  $$
  \mu_j \space for \space j\in [A,B,C,D] \\
  $$
```{r}
df = data.frame(obs = c(20,40,18,75,25,21,15,30,21,25,32,30), class = c(rep("A", 3), rep("B", 3), rep("C", 3), rep("D", 3)))
aov.fit = aov(obs ~ class, data = df)
summary(aov.fit)
```
_from the data summary above, we know the SSR is 2247, and the df of Residuals is 8. the P.value according to F test is 0.598._

##### b) Do the test for the contrast: H0 : (μA + μB)/2 = (μC + μD)/2.
  $$
  \begin{aligned}
  H_0: \frac{\mu_A + \mu_B}{2} &= \frac{\mu_C + \mu_D}{2} \\
       \frac{\mu_A}{2} + \frac{\mu_B}{2} - \frac{\mu_C}{2} - \frac{\mu_D}{2} &= 0 \\
       \sum_{j=1}^4 c_j &= 0 \\
  H_1: at \space least \space one \space not \space equal \\
       \sum_{j=1}^4 c_j &!= 0 \\
  \end{aligned}
  $$
  **method 1**
```{r}
Rss <- 2247
R.df <- 8
s.sq <- Rss / R.df
i <- 3
g <- 4
(data.table = matrix(data = c(20,40,18,75,25,21,15,30,21,25,32,30), nrow = i, ncol = g))
AvsBvsCvsD <- c(1/2,1/2,-1/2,-1/2)
var.thetaI.hat <- rep(NA, g)
thetaI.hat <-rep(NA, g)
for (j in 1:g) {
  length(data.table[,j])
  thetaI.hat[j] <- AvsBvsCvsD[j]*mean(data.table[,j])
  var.thetaI.hat[j] <- AvsBvsCvsD[j]^2 * s.sq / length(data.table[,j])
}
(theta.hat <- sum(thetaI.hat))
(var.theta.hat <- sum(var.thetaI.hat))
t <- theta.hat / sqrt(var.theta.hat)
(p.value <- 2 * pt(t, df = R.df, lower.tail = FALSE))
```

  **method 2**
```{r}
df = data.frame(obs = c(20,40,18,75,25,21,15,30,21,25,32,30), treatment = c(rep("A", 3), rep("B", 3), rep("C", 3), rep("D", 3)))
df$treatment = as.factor(df$treatment)
AvsB = c(1,-1,0,0)
AtoBvsCtoD = c(1/2,1/2,-1/2,-1/2)
Matrix = cbind(AvsB, AtoBvsCtoD)
Clist = list("AvsB" = 1, "AtoBvsCtoD" = 2)
contrasts(df$treatment) = Matrix
aov.model = aov(obs ~ treatment, data = df)
summary(aov.model,split=list(treatment=Clist))
```
_since the p-value of A to B vs. C to D is equal to 0.451, which is greater than the assumed level of significant value 0.05, therefore, we fail to reject H0._ 

## Question 3.
##### a) Test at the α = 0.10 level of significance the null hypothesis that viewing levels for news are the same for ABC, CBS, and NBC.
  $$
  H_0: \mu_{ABC} = \mu_{CBS} = \mu_{NBC} \\
  H_1: \mu_{ABC} \neq \mu_{CBS} \neq \mu_{NBC} \\
  $$
```{r}
alpha <- 0.10
news.dataset <- data.frame(data = c(19.7,18.6,19.1,17.9,16.1,15.8,14.7,18.1,18.2,17.9,14.8,19.2))
news.dataset$Broadcasts <- rep(c("ABC","CBS","NBC"), each = 4)
news.dataset$Cities <- rep(c("A","B","C","D"), 3)
news.dataset
summary(aov(data ~ Broadcasts+Cities, data = news.dataset))
```
_since the p-value for comparing the level of news in three broadcasts is 0.0762, which is less than the level of significant 0.1. Therefore, we reject the H0._


